Computing the {\em least-commitment} action model from state observations

### 2018-12-17 doc/ijcai-19-conformant/conformant-learning.tex
conformant-learning.adding comments by Eva and Diego from todays meeting
updates added after meeting for discussing the points made by blai

\section{Introduction}
Given a sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation. This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.

In addition, the paper introduces a new method to compute the {\em least-commitment} action model from a sequence of partially observed states. The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em conformant planning}.

\section{Background}
This section formalizes the planning models we use in the paper as well as the kind of observations that are given as input for the computation of the {\em least-commitment} action model.

\subsection{Classical planning with conditional effects}
Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables). A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. $L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents. A {\em state} $s$ is a full assignment of values to fluents. We explicitly include negative literals $\neg f$ in states s.t. $|s|=|F|$ and the size of the state space is $2^{|F|}$.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}. An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$. The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$. Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$. And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

@@

An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$. Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}. An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
\[
triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
\]

The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.

@@

A partially observable state $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable. Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable. Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$. In practice, the number of observed states $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.

In other words, we assume there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation). Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite. In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.

\begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
Given a {\em classical planning problem} $P$ and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that a plan $\pi$ {\em explains the observation} (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ that is consistent with $\mathcal{O}(\tau)$. If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$. 
\end{definition}

@@

\subsection{The {\em least-commitment} action model}
The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{\Phi,\mathcal{O}(\tau)}$:
\begin{itemize}
\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where $A[\cdot]$ is a set of actions s.t. the semantics of each $a\in A[\cdot]$ is unknown; i.e. the corresponding $\tup{\rho,\theta}$ functions are undefined. 
\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ that is defined within the {\em classical planning frame} $\Phi$.
\end{itemize}

Before formalizing the solution to this task, i.e. the {\em least-commitment} action model, we introduce several necessary definitions. Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$. We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$. Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.

Now, we define a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
\begin{definition}[Partially specified action model]
Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$. (Note that if $M$ is a singleton it represents a {\em fully specified action model}).
\end{definition}

@@

Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by constraints of three kinds:
\begin{enumerate}
\item {\em Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$. 
\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. As a rule of thumb, {\it state invariants} constraining the possible states of a given planning domain belong to this second class of constraints~\cite{fox1998automatic}. 
\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata. 
\end{enumerate}

\subsection{Computing the {\em least-commitment} model via conformant planning}
Given a task $\tup{\Phi,\mathcal{O}(\tau)}$, this section shows that we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$. In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G_c}$ such that:
\begin{itemize}
\item The set of fluents $F_c$ extends $F$ with two new sets of fluents:
      \begin{itemize}
      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
      \item Fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$. {\bf Our encoding exploits the syntactic constraint of \strips\ } so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$. Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
      \end {itemize}

@@

      \item The set of actions $A_c$ contains now actions of three different kinds:
\begin{itemize}
      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} fluents to a positive/negative value (similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} fluents to a positive/negative value).
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}\},\\
\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
                                    &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\\\
\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}\},\\
\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\},\\
                                    &\{\neg pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\}.                                    
\end{align*}
\end{small}

      \item Actions for {\em validating} that committed models explain the $s_j$ observed states, {\tt\small $0\leq j< m$}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
                            &\{mode_{commit}\}\rhd\{\neg mode_{commit}, mode_{val}\}.
\end{align*}
\end{small}

      \item {\em Editable} actions whose semantics is given by the value of {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state. Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema. Note that this editable schema when the set of fluents {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold at the current state, then it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}. Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
\begin{small}  
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
\cond(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi, eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
&\{\neg pre\_e\_\xi, eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
\end{align*}
\end{small}

@@

\item The clauses in $\Upsilon$ comprises:
      \begin{enumerate}
      \item The {\em unit clauses} given by the fluents that hold in the initial state $I=s_0$ and $mode_{commit}$ set to true.
      \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown. In other words, that any model from the \strips\ space of models (following the previously mentioned syntactic constraints) can initially be part of the {\em least-commitment} action model. Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
            \begin{itemize}
            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}.
            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}.
            \end{itemize}
      \end{enumerate}
One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier for a specific domain. 
\item The new goals are $G_c=s_m\cup\{test_m\}$.
\end{itemize}

>>>>

\section{Introduction}
Given an input sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation. This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.

In addition, the paper introduces a new method to compute the {\em least-commitment} action model for an input sequence of partially observed states. The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em conformant planning}.

\section{Background}
This section formalizes the {\em planning models} we use in the paper as well as the kind of state {\em observations} that are given as input for computing the {\em least-commitment} action model.

\subsection{Classical planning with conditional effects}
Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values). Given $L$, let $\neg L=\{\neg l:l\in L\}$ be its complement. We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents. A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}. Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$. The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$. Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$. The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$. Subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state. The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.

@@

An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$. Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}. An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\neg\eff_c(s,a)\cup\eff_c(s,a)\}$ where $\eff_c(s,a)$ are the {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$):
\[
\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
\]

@@

A {\em partial state} $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable. Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable. Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be {\em consistent} with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$. In practice, the number of observed states $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.

\newpage

We are assuming then that there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation). Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite. In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.

\begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
Given a {\em classical planning problem} $P$ and a sequence of partial states $\mathcal{O}(\tau)$, a plan $\pi$ {\em explains} $\mathcal{O}(\tau)$ (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ {\em consistent} with $\mathcal{O}(\tau)$.  
\end{definition}
If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.

Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$. We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$. Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.

@@

\subsection{The {\em least-commitment} action model}
The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{\Phi,\mathcal{O}(\tau)}$:
\begin{itemize}
\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e. the corresponding $\tup{\rho,\theta}$ functions are undefined. 
\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ defined within the {\em classical planning frame} $\Phi$.
\end{itemize}

Before formalizing the solution to this task, i.e. the {\em least-commitment} action model, we introduce several necessary definitions.

\begin{definition}[Model Space]
Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ the model space $M$ is the set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$. 
\end{definition}

Now, we define a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
\begin{definition}[Partially specified action model]
A {\em partially specified action model} is a subset of models in a given {\em model space $M$}.
\end{definition}
If the {\em partially specified action model} is a singleton, it represents a {\em fully specified action model}. On the other hand, if its size is $|M|$ the {\em partially specified action model} represents the full model space.

@@

Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by constraints of three kinds:
\begin{enumerate}
\item {\em Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$. With this regard, we implement a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$ using only fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$). This encoding exploits the syntactic constraints of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$. Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. As a rule of thumb, {\em typing constraints} and {\it state invariants} belong to this second class of constraints~\cite{fox1998automatic}. 
\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata. For instance, since we assume that the initial state $s_0$ is fully observed, we have that $\bigcup_{\xi}$ {\tt\small $\neg\pre\_e\_\xi$} $\wedge$ {\tt\small $\eff\_e\_\xi$} if there is an instantiation of $e$ that holds in $s_i^o$ ({\small $0<i$}) but not in $s_0$.

\end{enumerate}

@@

\subsection{Computing the {\em least-commitment} model via classical planning}
Inspired by the {\em classical planning compilation} $K_{s_0}$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$: 
\begin{itemize}
\item The set of fluents $F'$ extends $F$ with two new sets of fluents:
      \begin{itemize}
      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models (note that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding).
      \end {itemize}
\item The set of actions $A'$ contains now actions of three different kinds:
\begin{itemize}
      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} to a positive/negative value. Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} but only to a positive value (recall that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding). The value of an effect {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding precondition {\tt\small pre\_e\_$\xi$} is committed (i.e. once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ hold in the current state).
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi}, \mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\pre\_e\_\xi}\}.\\\\
\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\pre\_e\_\xi},\mathsf{\neg K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.\\\\
\hspace*{7pt}\pre(\mathsf{commit\top\_eff\_e\_\xi})=&\{mode_{commit}, \\
&\mathsf{\neg K\eff\_e\_\xi},\\
&\mathsf{K\pre\_e\_\xi}\vee \mathsf{K\neg\pre\_e\_\xi}\},\\
\cond(\mathsf{commit\top\_eff\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\eff\_e\_\xi}\}.\\\\
\end{align*}
\end{small}

      \item Actions for {\em validating} that committed models explain the $s_j$ observed states, {\tt\small $0\leq j< m$}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\},\\
                            &\{mode_{commit}\}\rhd\{\neg mode_{commit}, mode_{val}\}.
\end{align*}
\end{small}

      \item {\em Editable} actions whose semantics is given by the value of fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} at the current state. Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema. This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state. Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
\begin{small}  
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
&\{K\neg\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
\end{align*}
\end{small}

@@

\item The new initial state $I'=I \cup\{mode_{commit}\}$ while the new goals are $G'=s_m\cup\{test_m\}$.
\end{itemize}

\begin{figure}
  \begin{tiny}  
  \begin{verbatim}
(:action stack
 :parameters (?o1 - object ?o2 - object)
 :precondition
   (and (or (not (Kpre_on_v1_v1_stack)) (on ?o1 ?o1))
        (or (not (Kpre_on_v1_v2_stack)) (on ?o1 ?o2))
        (or (not (Kpre_on_v2_v1_stack)) (on ?o2 ?o1))
        (or (not (Kpre_on_v2_v2_stack)) (on ?o2 ?o2))
        (or (not (Kpre_ontable_v1_stack)) (ontable ?o1))
        (or (not (Kpre_ontable_v2_stack)) (ontable ?o2))
        (or (not (Kpre_clear_v1_stack)) (clear ?o1))
        (or (not (Kpre_clear_v2_stack)) (clear ?o2))
        (or (not (Kpre_holding_v1_stack)) (holding ?o1))
        (or (not (Kpre_holding_v2_stack)) (holding ?o2))
        (or (not (Kpre_handempty_stack)) (handempty)))
 :effect (and
   (when (and (Kpre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
   (when (and (Kpre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
   (when (and (Kpre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
   (when (and (Kpre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
   (when (and (Kpre_ontable_v1_stack)(Keff_ontable_v1_stack)) (not (ontable ?o1)))
   (when (and (Kpre_ontable_v2_stack)(Keff_ontable_v2_stack)) (not (ontable ?o2)))
   (when (and (Kpre_clear_v1_stack)(Keff_clear_v1_stack)) (not (clear ?o1)))
   (when (and (Kpre_clear_v2_stack)(Keff_clear_v2_stack)) (not (clear ?o2)))
   (when (and (Kpre_holding_v1_stack)(Keff_holding_v1_stack)) (not (holding ?o1)))
   (when (and (Kpre_holding_v2_stack)(Keff_holding_v2_stack)) (not (holding ?o2)))
   (when (and (Kpre_handempty_stack)(Keff_handempty_stack)) (not (handempty)))
   (when (and (Knot_pre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (on ?o1 ?o1))
   (when (and (Knot_pre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (on ?o1 ?o2))
   (when (and (Knot_pre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (on ?o2 ?o1))
   (when (and (Knot_pre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (on ?o2 ?o2))
   (when (and (Knot_pre_ontable_v1_stack)(Keff_ontable_v1_stack)) (ontable ?o1))
   (when (and (Knot_pre_ontable_v2_stack)(Keff_ontable_v2_stack)) (ontable ?o2))
   (when (and (Knot_pre_clear_v1_stack)(Keff_clear_v1_stack)) (clear ?o1))
   (when (and (Knot_pre_clear_v2_stack)(Keff_clear_v2_stack)) (clear ?o2))
   (when (and (Knot_pre_holding_v1_stack)(Keff_holding_v1_stack)) (holding ?o1))
   (when (and (Knot_pre_holding_v2_stack)(Keff_holding_v2_stack)) (holding ?o2))
   (when (and (Knot_pre_handempty_stack)(Keff_handempty_stack)) (handempty))))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small PDDL encoding of the editable version of the {\tt\small stack(?v1,?v2)} schema.}
\label{fig:editable}
\end{figure}
